# orbis <a href = "https://danielvartan.github.io/brandr/"><img src = "man/figures/logo.png" align="right" width="120" /></a>

<!-- quarto render -->

```{r}
#| label: setup
#| include: false

library(orbis)

library(badger)
library(beepr)
library(cffr)
library(codemetar)
library(groomr) # https://github.com/danielvartan/groomr
library(here)
library(rutils) # https://github.com/danielvartan/rutils
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_repostatus("active"),
  badger::badge_github_actions(action = "R-CMD-check.yaml"),
  badger::badge_codecov() |> rutils::shush(),
  badger::badge_license(
    "MIT",
    "green",
    url = "https://choosealicense.com/licenses/mit/"
  ),
  badger::badge_custom(
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```
<!-- badges: end -->

## Overview

`orbis` is an R package that offers a comprehensive suite of tools for simplifying spatial data analysis.

## Installation

You can install `orbis` using the [`remotes`](https://github.com/r-lib/remotes) package:

``` r
remotes::install_github("danielvartan/orbis")
```

## Usage

`orbis` is equipped with several functions to help with your analysis, such as:

- [`filter_points_on_land`](https://danielvartan.github.io/orbis/reference/Filter_points_on_land.html): filters latitude/longitude points that intersects with a given [`sf`](https://r-spatial.github.io/sf/) geometry.
- [`get_country_names`](https://danielvartan.github.io/orbis/reference/get_country_names.html): Get names or codes of all countries present in the [ISO 3166-1](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3) standard.
- [`get_brazil_region`](https://danielvartan.github.io/orbis/reference/get_brazil_region.html): Get Brazilian regions.
- [`get_brazil_fu`](https://danielvartan.github.io/orbis/reference/get_brazil_fu.html): Get Brazilian federative unit abbreviations.
- [`get_brazil_state`](https://danielvartan.github.io/orbis/reference/get_brazil_state.html): Get Brazilian state names.
- [`get_brazil_municipality`](https://danielvartan.github.io/orbis/reference/get_brazil_municipality.html): Get Brazilian municipalities data.
- [`get_sidra_by_year`](https://danielvartan.github.io/orbis/reference/get_sidra_by_year.html): Get and aggregate data by year from [SIDRA](https://sidra.ibge.gov.br/) API (to avoid overloading).
- [`wc_to_ascii`](https://danielvartan.github.io/orbis/reference/wc_to_ascii.html): Convert [WorldClim](https://worldclim.org/) [GeoTIFF](https://en.wikipedia.org/wiki/GeoTIFF) files to [ASCII](https://en.wikipedia.org/wiki/Esri_grid) raster format.

Click [here](https://danielvartan.github.io/orbis/) to see the full list.

## License

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_license(
    "MIT",
    "green",
    url = "https://choosealicense.com/licenses/mit/"
  )
)
```

`orbis` code is released under the [MIT license](https://opensource.org/license/mit).

## Contributing

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_custom( # Contributor Covenant
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```

Contributions are welcome, including bug reports. Take a moment to review our [Guidelines for Contributing](https://danielvartan.github.io/orbis/CONTRIBUTING.html).

[![](https://img.shields.io/static/v1?label=Sponsor&message=%E2%9D%A4&logo=GitHub&color=%23fe8e86)](https://github.com/sponsors/danielvartan)

You can also support the development of `orbis` by becoming a sponsor. Click [here](https://github.com/sponsors/danielvartan) to make a donation. Please mention `orbis` in your donation message.
