# orbis <a href = "https://danielvartan.github.io/brandr/"><img src = "man/figures/logo.png" align="right" width="120" /></a>

<!-- quarto render -->

```{r}
#| label: setup
#| include: false

library(orbis)

library(badger)
library(beepr)
library(cffr)
library(codemetar)
library(dplyr)
library(fs)
library(geodata)
library(groomr) # https://github.com/danielvartan/groomr
library(here)
library(readr)
library(rutils) # https://github.com/danielvartan/rutils
library(terra)
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_repostatus("active"),
  badger::badge_cran_release(),
  badger::badge_github_actions(action = "R-CMD-check.yaml"),
  badger::badge_codecov() |> rutils::shush(),
  badger::badge_license(
    "MIT",
    "green",
    url = "https://choosealicense.com/licenses/mit/"
  ),
  badger::badge_custom(
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```
<!-- badges: end -->

## Overview

`orbis` is an R package that offers a suite of tools for simplifying spatial data analysis.

> If you find this project useful, please consider giving it a star! &nbsp; [![GitHub repo stars](https://img.shields.io/github/stars/danielvartan/orbis)](https://github.com/danielvartan/orbis/)

## Installation

You can install `orbis` using the [`remotes`](https://github.com/r-lib/remotes) package:

``` r
remotes::install_github("danielvartan/orbis")
```

## Usage

`orbis` is equipped with several functions to help with your analysis, such as:

- [`filter_points_on_land`](https://danielvartan.github.io/orbis/reference/Filter_points_on_land.html): Filters latitude/longitude points that intersects with a given [`sf`](https://r-spatial.github.io/sf/) geometry.
- [`get_brazil_municipality`](https://danielvartan.github.io/orbis/reference/get_brazil_municipality.html): Get Brazilian municipalities data.
- [`get_sidra_by_year`](https://danielvartan.github.io/orbis/reference/get_sidra_by_year.html): Get and aggregate data by year from [SIDRA](https://sidra.ibge.gov.br/) API (to avoid overloading).
- [`shift_and_rotate`](https://danielvartan.github.io/orbis/reference/shift_and_rotate.html): Shift and rotate raster or vector data.
- [`shift and_crop`](https://danielvartan.github.io/orbis/reference/shift_and_crop.html): Shift, rotate, and crop a raster using a vector.
- [`wc_to_ascii`](https://danielvartan.github.io/orbis/reference/wc_to_ascii.html): Convert [WorldClim](https://worldclim.org/) [GeoTIFF](https://en.wikipedia.org/wiki/GeoTIFF) files to [ASCII](https://en.wikipedia.org/wiki/Esri_grid) raster format.

Example:

```{r}
#| eval: false

library(dplyr)
library(geodata)
library(orbis)
library(terra)
```

```{r}
#| label: readme-shift-and-crop-1

raster <-
  expand.grid(
    seq(-179.75, 179.75, by = 0.5),
    seq(-89.75, 89.75, by = 0.5)
  ) |>
  as_tibble() |>
  rename(x = Var1, y = Var2) |>
  mutate(value = rnorm(259200)) |>
  rast(type = "xyz") %>%
  `crs<-`("epsg:4326")

world_shape <- world(path = tempdir())
raster <- raster |> crop(world_shape, mask = TRUE)

raster |> plot()
```

```{r}
#| label: readme-shift-and-crop-2

vector <- gadm(country = "rus", level = 0, path = tempdir())

vector |> plot()
```

```{r}
#| label: readme-shift-and-crop-3

raster |>
  shift_and_crop(vector, dx = -45) |>
  plot()
```

Click [here](https://danielvartan.github.io/orbis/reference/) to see the full list of functions.

## License

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_license(
    "MIT",
    "green",
    url = "https://choosealicense.com/licenses/mit/"
  )
)
```

`orbis` code is released under the [MIT license](https://opensource.org/license/mit). This means you can use, modify, and distribute the code as long as you include the original license in any copies of the software that you distribute.

## Contributing

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_custom( # Contributor Covenant
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```

Contributions are welcome, including bug reports. Take a moment to review our [Guidelines for Contributing](https://danielvartan.github.io/orbis/CONTRIBUTING.html).

[![](https://img.shields.io/static/v1?label=Sponsor&message=%E2%9D%A4&logo=GitHub&color=%23fe8e86)](https://github.com/sponsors/danielvartan)

You can also support the development of `orbis` by becoming a sponsor. Click [here](https://github.com/sponsors/danielvartan) to make a donation. Please mention `orbis` in your donation message.
